Improving legitimacy and efficiency \\ of decisions in democratic communities

### 2017-09-19 statistical-quorum.tex
Methodology section done.

\begin{abstract}
The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
maximize productivity?
We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to focus
attention, in turn making the entire decision-making process dynamic.
We developed an online system centered around this metric and tested it in a 140-person community that holds weekly
councils over the course of 16 weeks.  We found that the community was able to use the online system to simultaneously
improve the legitimacy and efficiency of their decision-making process.
\end{abstract}

@@

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results legitimate.
\item We implemented an online application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases legitimacy of decisions,
improves utilization of collective attention, increases overall participation,
and increases number of decisions made while not sacrificing their legitimacy.
\end{itemize}

@@

To evaluate statistical quorum in practice, we developed an online web application PeerMind and
used it in a democratically run community of 140 members for 16 weeks, replacing the traditional
decision making through in-person meetings. The community is a house of 140
students living together and making collective decisions about the governance and operation of
the house, use of resources like common spaces and budgets, and changes to the house itself.

@@

To evaluate general perception of members before and after the use of the application for decision making,
we did a general survey at the end of each semester, asking members to answer general question about how
they felt the decision making in the house was that semester.

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
of the community were invited to participate in the survey.}
\label{fig:survey-comparison}
\end{figure}

@@

@@

\todo[inline]{See my comment at the beginning of the PeerMind Web App section.  I think the evaluation you note below about the interviews is out of the scope of this paper, which should be focused on statistical quorum and less online vs. offline platforms.\\
---Max}

>>>>

\begin{abstract}
The Achilles heel of many major democracies is the eternal need for sufficient engagement and attention from the community.
While there have been major efforts towards increasing these attributes within the community, we explore a fundamentally
different approach: For a fixed level of community attention, how can we optimize the allocation of this attention to
maximize productivity?
We propose a novel \emph{statistical quorum}, derived using first principles, as one metric the community can use to focus
attention, in turn making the entire decision-making process dynamic.
We developed a web application centered around this metric and tested it in a 140-person community that holds weekly
councils over the course of 16 weeks.  We found that the community was able to use the web application to simultaneously
improve the legitimacy and efficiency of their decision-making process.
\end{abstract}

@@

\begin{itemize}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed a novel concept of \emph{statistical quorum}
as a threshold of votes required to deem results legitimate.
\item We implemented a web application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in a decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases legitimacy of decisions,
improves utilization of collective attention, increases overall participation,
and increases number of decisions made while not sacrificing their legitimacy.
\end{itemize}

@@

To evaluate statistical quorum in practice, we developed a web application PeerMind and
used it in a democratically run community of 140 members for 16 weeks, replacing the traditional
decision making through in-person meetings. The community is a house of 140
students living together and making collective decisions about the governance and operation of
the house, use of resources like common spaces and budgets, and changes to the house itself.

@@

To evaluate general perception of members before and after the use of the application for decision making,
we did a general survey at the end of each semester, asking members to answer general questions about the
decision making in the house that semester.

We instrumented the application to collect information on how much time users spent using the application,
measuring the time the application was focused. Other information we used for evaluation are available
through regular features of the application: how many and who made discussion items, comments, motions
(proposed decisions on discussion items), and vote tallies per motion.

For the baseline, we extracted similar information from minutes: how long did the councils last, how many
discussions were opened, how many comments by which members were made, how many motions were proposed,
and vote tallies for them.

Finally, we did semi-structured interviews at the end of the fall semester with 7 members who were participating
both in spring and fall semesters and ask them to compare their experience between those two semesters and
evaluate the application and the use of statistical quorum.

@@

\begin{figure}[ht]
\centering
\includegraphics[width=0.5\textwidth]{figures/survey-comparison.pdf}
\caption{Results of a survey at the end of spring and fall semesters about general perception of the decision making (councils) in the community.
For each question, a participant was asked to rank on a scale between 1 (completely disagree) and 7 (completely agree). All members
of the community were invited to participate in the survey each time.}
\label{fig:survey-comparison}
\end{figure}

@@
